\Title

### 2021-01-26 manuscript/manuscript.tex
Improvements to conclusions

\section{Gridding gravity data from Australia}

@@

This section will demonstrate how gradient-boosted equivalent sources can be
used to interpolate large datasets of gravity data onto a regular grid at
uniform height.
For this purpose, we selected a compilation of ground gravity surveys over
Australia made by \citet{wynne2018} (TODO: MENTION THE COMPGEOLAB VERSION?).
It contains over 1.7 million data points and covers most of the Australian
territory at variable point spacings.
Our goal is to create a 1 arc-minute resolution grid of gravity disturbances at
a constant geometric height of \AustraliaEqlGridHeight{} (the largest height of
observations).

We computed the gravity disturbance by removing the normal gravity of
the WGS84 ellipsoid from the observed gravity data (Fig.~\ref{fig:australia}).
Here, normal gravity was computed at each observation point through the
closed-form formula of \citet{ligotze2001} (TODO: ADD CITATION TO BOULE).
Finally, we convert the observations to planar Cartesian coordinates by
applying a Mercator projection.

We start the interpolation process by defining a set of block-averaged sources
at a relative depth of \AustraliaEqlDepth{} and a block size of
\SI{1.8}{\kilo\meter}, resulting in a total of \AustraliaEqlNSources{}~point
sources.
The block size was chosen to match the desired resolution of the final grid (1
arc-minute is approximately \SI{1.8}{\kilo\meter} at the equator).
The coefficients of the gradient-boosted equivalent sources are estimated using
a damping equal to \AustraliaEqlDamping{} and a window size of
\AustraliaEqlWindowSize{}, which was chosen in order to limit the amount of
computer memory needed to store the Jacobian matrix to under 10 Gigabytes.
Finally, we used estimated source coefficients to predict the values of the
gravity disturbance on a regular grid of
\AustraliaEqlGridNLongitude{}$\times$\AustraliaEqlGridNLatitude{} points at
\AustraliaEqlGridHeight{} above the ellipsoid.
Fig.~\ref{fig:australia} shows the original data distribution and the
interpolated regular grid.
Grid points that are further than \SI{50}{\kilo\meter} from the nearest data
point are masked to avoid unrealistic extrapolations.

The resulting grid preserves the high resolution of the original data while
avoiding aliasing artifacts because to the block averaging.
Some parts of the grid are smoother and with reduced amplitude (e.g., some
southwestern parts), which is expected from the upward continuation that was
performed to have the grid at a constant height.
On a modest workstation with 16 cores and 16 Gigabytes of RAM,
estimating the \AustraliaEqlNSources{} coefficients with gradient-boosting took
$\sim 1.3$ hours and the forward modelling step took $\sim 18$ minutes.

FOR DISCUSSION:

If the grid will have a much higher resolution than the block spacing, the
equivalent sources will not be able to reproduce the high frequency components
of the data.
Therefore we recommend setting a block spacing that matches the resolution of
the final grid.

@@

>>>>

\section{Gridding Australia gravity data}

@@

Our final goal is to show how gradient-boosted equivalent sources can be used
to interpolate a very large dataset of gravity data into a regular grid.
We selected a compilation of gravity surveys over Australia made by
\citet{wynne2018}.
It consists in more than 1.7 million data points that cover most part of the
Australia territory obtained through observations on ground.
First we compute the gravity disturbance by removing the normal gravity from
the observed gravity data (Fig.~\ref{fig:australia}).
The normal gravity is computed on each observation point through the
closed-form formula given by \citet{ligotze2001} and using the WGS84 ellipsoid
as reference.
Then we obtain planar coordinates of the observation points by applying
a Mercator projection.

We start the interpolation process by defining a set of block-averaged sources
at a relative depth of \AustraliaEqlDepth{} and a block spacing of
\SI{1.8}{\kilo\meter}.
These parameters create \AustraliaEqlNSources{}~point sources.
The choice of the block spacing must be done taking into account the desired
resolution of the final grid.
If the grid will have a much higher resolution than the block spacing, the
equivalent sources will not be able to reproduce the high frequency components
of the data.
Therefore we recommend setting a block spacing that matches the resolution of
the final grid.
In this case we will generate a regular grid with a resolution of 1~arc-minute,
which can be approximated by \SI{1.8}{\kilo\meter} for the Earth.

The coefficients of the gradient-boosted equivalent sources are estimated using
a damping equal to \AustraliaEqlDamping{} and a window size of
\AustraliaEqlWindowSize{} in order to limit the amount of memory needed to
store the largest Jacobian matrix to 10GB.

Finally, the equivalent sources are used to predict the values of the gravity
disturbance on a regular grid of
\AustraliaEqlGridNLongitude{}$\times$\AustraliaEqlGridNLatitude{} points at
\AustraliaEqlGridHeight{} above the ellipsoid
(Fig.~\ref{fig:australia}).

@@

The equivalent source technique has been proven to be a well suited method for
interpolating any harmonic data such as gravity disturbances and magnetic
anomalies.
The two main reasons that make it to stand out over all-purpose 2D
interpolation methods is the fact that the equivalent sources take into account
the height of the observations and they always produce harmonic predictions.
The main inconvenience of using equivalent sources is the computational
requirements needed to estimate the coefficients of the equivalent sources,
specially on the computer memory needed to store the Jacobian matrix.
This forbids the application of the equivalent source technique to very
large datasets where the computer memory needed to store the whole Jacobian
matrix greatly exceeds the specifications of modest personal computers now days.

In this work we present two strategies that could be simultaneously applied in
order to interpolate very large datasets in reasonable times without the need
to store large matrices into memory:
the block-averaged sources, which reduce the number of
equivalent sources needed for the interpolation;
and the gradient-boosted equivalent sources that simplify the
problem of estimating the source coefficients into smaller problems
through sets of equivalent sources defined by overlapping windows.

Both methods were exhaustively tested against a synthetic model and synthetic
surveys in order to compare accurate they are in comparison to the existing
techniques and how they perform in terms of computational resources.
Block-averaged sources were compared to the interpolations obtained with two
traditional strategies for locating the sources:
the \emph{source below data} and the \emph{regular grid} of sources.
The performance of the gradient-boosted equivalent sources was compared against
the regular non-boosted equivalent sources, along with tests for assessing how
the choice of the window size and the amount of overlap between adjacent
windows affect both the accuracy of the interpolations and the computational
load.

Our results showed that the block-averaged sources reduce the computational
load needed to estimate their coefficients in comparison to the two traditional
strategies mentioned before.
They also showed that this reduction of the number of sources doesn't affect
the accuracy of the predictions, making block-averaged sources a better
alternative to the traditional methods.
Besides, block-averaged sources might also prevent aliasing on the interpolated
values, specially when the observation points show high anisotropy, i.e. they
were taken along paths with a high sample frequency.
From our experimentation with this new strategy we noticed that special
attention must be payed when choosing the size of the block spacing:
as a thumb rule, it must be approximately equal to the resolution of the
regular grid where the values will be interpolated.

We have also carried out similar tests to compare the vertical location of the
sources.
Our results showed that any one of the three strategies taken into account
(\emph{constant depth}, \emph{relative depth} and \emph{variable depth})
doesn't significantly affects the accuracy of the interpolations.
Nevertheless, we are more prone to choose either the \emph{constant depth} or
the \emph{relative depth} for any application because they involve less
hyper-parameters that would be needed to estimated before the actual
interpolation.

Regarding the gradient-boosted equivalent sources, our results show that they
heavily reduce the computer memory needed to estimate source coefficients,
making possible to interpolate even very large datasets.
The interpolations obtained though this new method achieve almost the same
accuracy than the non-boosted sources, while reducing the computation time
needed to estimate the source coefficients.
Because it doesn't involve any assumption on the location of the sources, it
can be used in conjunction with any horizontal source layout and any strategy
for setting their depth.
This allows to use both the block-averaged sources and the gradient-boosted
equivalent sources simultaneously.
Our results allowed us to develop a rule of thumb in order to maximize the
accuracy of the interpolations: the size of the overlapping windows must be
chosen as the maximum value that creates Jacobian matrices that fit into our
computer memory.
Therefore its choice is subject to the particular infrastructure where the
interpolation will be carried out.
We also show that an overlap of 50\% between adjacent windows achieves a good
compromise between accuracy and computation time needed to fit the source
coefficients.

Finally, both proposed methods were used to interpolate a compilation of more
than 1.7 million gravity points from Australia on a regular grid with
a resolution of 1~arc-minute.
The whole process was carried out on a modest personal computer, requiring
a maximum of 10GB of memory.
This shows how the new gradient-boosted equivalent sources can be applied to
real-world data.
